### 2021-04-13 introduction.tex
update contributions

We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing process calculi. Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi. Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types. In concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates. In process calculi, these are conflated. Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be embedded in CP~\cite{wadler14}, and consequently it guarantees the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
Following the above observations, we thus pose our research question for GV.

@@

We address this research question by following the line of work from CP to Priority CP, and present
Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
In this paper we make the following contributions:
\begin{enumerate}
\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate*}
	\item our version does not have the pseudo-type $S^\sharp$;
	\item our structural congruence is type preserving.
	\end{enumerate*}
\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
\item \textbf{Connection to linear logic}. The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}). We recall Milner's cyclic scheduler~\cite{milner89} presented in PCP~\cite{dardhagay18} and encode it in PGV.
\end{enumerate}
\end{document}

>>>>

We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing process calculi. Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi. Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types. In concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates. In process calculi, these are conflated. Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be translated into CP~\cite{wadler14}, thus guaranteeing the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
Following the above observations, we thus pose our research question for GV.

@@

We address this research question by following the line of work from CP to Priority CP, and present
Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
%
In this paper we make the following main contributions:
\begin{enumerate}
	\item \textbf{Priority GV}. We present Priority GV (PGV), a session-typed functional language with priorities (\cref{sec:pgv}) and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
	PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate*}
	\item we do not require the pseudo-type $S^\sharp$; and
	\item our structural congruence is type preserving.
	\end{enumerate*}
	PGV answers our research question as it allows cyclic-structured binary session-typed processes that are deadlock free.
	\item \textbf{Updated PCP}. We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
	\item \textbf{Translation of PGV to PCP}. We present a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
\end{enumerate}

%\begin{enumerate}
%\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
%PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
%\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
% 	\begin{enumerate*}
%	\item our version does not have the pseudo-type $S^\sharp$;
%	\item our structural congruence is type preserving.
%	\end{enumerate*}
%\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
%\item \textbf{Connection to linear logic}. The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}). We recall Milner's cyclic scheduler~\cite{milner89} presented in PCP~\cite{dardhagay18} and encode it in PGV.
%\end{enumerate}
\end{document}
